---
title: "Homework Assignment 2"
author: "Anna Sehgal"
format:
  html:
    toc: true
    toc-depth: 3
    number-sections: true
    code-fold: false
execute:
  echo: true
  warning: false
  message: false
---

## March work day (10 pts)

R functions for common set operations (when you have *actual sets/vectors* of outcomes):

```{r, echo = TRUE, eval = FALSE}
x <- c(1, 2, 3)
y <- c(2, 3, 4, 5)
union(x, y)      # combined elements
intersect(x, y)  # shared elements
setdiff(x, y)    # elements in x but not y (order matters)
setequal(x, y)   # are the sets equal?
```

You have a vector of March 2026 dates (`march`), and vectors `weekends` and `springbreak`. Use R’s set operations to find the dates of **non‑vacation weekdays**.

```{r}
march <- 1:31
march
weekends <- c(1, 7, 8, 14, 15, 21, 22, 28, 29)
weekends
springbreak <- 16:20
springbreak
all_holidays <- union(weekends, springbreak)
all_holidays
non_vacation_weekdays <- setdiff(march, all_holidays)
non_vacation_weekdays
```
```{r}
total_non_vacation_weekdays <- length(non_vacation_weekdays)
total_non_vacation_weekdays 
```

## Two yellow balloons (20 pts)

**Problem :**  
There are 16 balloons: 5 yellow, 8 red, 3 green. Two darts pop two balloons (without replacement).

Let:

- $P(1_{yellow})$ = probability first balloon is yellow
- $P(2_{yellow})$ = probability second balloon is yellow

1. Given first is yellow, find $P(2_{yellow}\mid 1_{yellow})$.


First balloon (yellow) is popped then:

$5-1 = 4$ (since first is popped, so we minus one)

$16-1 = 15$ (since one is popped and not replaceable, we decrease one from total)

So $P(2_{yellow}\mid 1_{yellow}) = \frac{4}{15} = 0.2666667$
```{r}
4/15
```


2. Find $P(1_{yellow}\cap 2_{yellow})$ (both balloons yellow).

$$P(A \cap B) = P(A) \times P(B|A),$$
$P(A) = P(1_{yellow})$ and $P(B|A) = P(2_{yellow}\mid 1_{yellow})$

$$P(1_{yellow} \cap 2_{yellow}) = P(1_{yellow}) \times P(2_{yellow}|1_{yellow})$$
$P(A) = P(1_{yellow})$ = $5/16$ = $0.3125$ (first balloon being yellow)
```{r}
5/16
```

$P(B|A) = P(2_{yellow}\mid 1_{yellow})$ = $4/15$ = $0.2666667$ (probablity of second being yellow that is popped)
```{r}
4/15
```

from question 1: $P(2_{yellow}\mid 1_{yellow}) = \frac{4}{15}  = 0.2666667$

And finally for both is:

$P(1_{yellow} \cap 2_{yellow}) = P(1_{yellow}) \times P(2_{yellow}|1_{yellow})$ = $0.3125 \times 0.2666667$
```{r}
0.3125*0.2666667
```
$P(1_{yellow} \cap 2_{yellow})$ = $0.08333334$


## Sequential system with A, B, C (15 pts)

A sequential system fails if **any** component fails. Components fail independently with:

- $p_A=0.02$
- $p_B=0.01$
- $p_C=0.03$

Find $P(\text{system fails})$.

Fail for independent event probablity:

P(A) = 0.02

P(B) = 0.01

P(C) = 0.03


System works
$A^C: 1 - 0.02 = 0.98$

```{r}
1 - 0.02
```

$B^C: 1 - 0.01 =0.99 $

```{r}
1 - 0.01
```

$C^C : 1 - 0.03 = 0.97$
```{r}
1 - 0.03
```

P(Working system) = $(P(A^C \cap B^C \cap C^C))$ = $(P(A^C) * P(B^C) * P(C^C))$ = $(0.98 \times 0.99 \times 0.97)$ = $0.941094$
```{r}
system_working <- 0.98*0.99*0.97
system_working
```

Probablity of System failure : 1 - P(Working system) = $1 - 0.941094$ = $0.058906$
```{r}
system_failure <- 1 - 0.941094
system_failure
```


## Problem parallel system (15 pts) 

A parallel system functions if **at least one** component functions. Components function **independently** with:

- $p_D=0.78$
- $p_E=0.75$
- $p_F=0.82$

Find $P(\text{parallel system functions})$.

P(D) = 0.78

P(E) = 0.75

P(F) = 0.82


Case if atleast one functions fails: 

$D^C: 1 - 0.78 = 0.22$
```{r}
1 - 0.78
```

$E^C: 1 - 0.75 = 0.25$
```{r}
1 - 0.75
```

$F^C: 1 - 0.82 = 0.18$
```{r}
1 - 0.82
```

P(work system) = $1- (P(D^C \cap E^C \cap F^C))$ = $1 - (P(D^C) * P(E^C) * P(F^C))$ = $(0.22 \times 0.25 \times 0.18)$ = $1 - 0.0099$ = $0.9901$
```{r}
system_fails <- 0.22*0.25*0.18
system_fails
```

```{r}
system_working_atleastone <- 1 - 0.0099
system_working_atleastone
```

## Spam detection using keywords problem (Bayes' Theorem) (20 pts)


An email service provider is designing a spam filter. From historical data, the company knows:

* **20%** of all incoming emails are spam.
* A specific keyword (for example, “FREE!!!”) appears in:

  * **60%** of spam emails
  * **5%** of legitimate (non-spam) emails

Suppose a new email arrives and it **contains this keyword**. What is the probability that the email is spam? 

To do: Find if it contains keyword and spam $P(S|KE)$

Let S = all emails are spam, KE = keyword 

$P(S) = 20% = 0.20$

Non-spam = $P(S^C)$ = $1- 0.20$ = 0.8
```{r}
non_spam <- 1- 0.20
non_spam
```

$P(KE | S) = 60% = 0.60 $ (keyword in spam)

$P(KE |S^C) = 5% = 0.05 $ (keyword is legitimate)

Formula used: $$P(B|A) = {P(A \cap B) \over P(A)} = {P(A|B) \times P(B) \over P(A|B) \times P(B) + P(A|\bar B) \times P(\bar B)}$$

So: $P(S|KE) = {P(KE|S) \times P(S) \over P(KE| S) \times P(S) + P(KE|\bar S) \times P(\bar S)}$

$=> P(S|KE) = {0.60 \times 0.20 \over 0.60 \times 0.20 + 0.05 \times 0.8}$ = 0.75 = 75% is spam having a keyword
```{r}
p_of_email_being_spam <- (0.60*0.20)/((0.60*0.20) + (0.05* 0.8))
p_of_email_being_spam
```



## Two way contingency table (20 pts)

Suppose you are helping the police develop a new test for blood alcohol levels. The potential advantage of the test is that it can be used during a routine traffic stop, but neither you nor the police are sure how accurate the test is. By design, the test should be positive when blood alcohol level is above $0.05\%.$

The test has been used on a large number of subjects where a more expensive, less convenient test is known to give perfectly accurate results. The table below shows the proportions for the outcome of the new test and true blood alcohol status:

| Test result | Alcohol level > 0.05%  | Alcohol level <= 0.05% |
|-------------|------------------------|------------------------|
| Positive    | 0.08                   | 0.3                    |
| Negative    | 0.02                   | 0.6                    |
	
a) What is the probability that a randomly selected driver has both a positive test and a blood alcohol level higher than $0.05\%$? 

- Let A = positive test, B = alcohol level > 0.05%
- Using formula $P(A \cap B)$ since it is an intersection in this case.

***= P(Alcohol level > 0.05% $\cap$ Positive) = 0.08 = 8%***

b) For a randomly selected driver, what is the probability that the test will be positive? 

- P(A) = sum of positive tests
$$P(A) = P(A \cap B) + P(A \cap \bar B) $$ = 0.08 + 0.3 = 0.38 = 38%

```{r}
positive_probablity <- 0.08 + 0.3
positive_probablity
```

	
c) Among drivers who have a positive test, what is the probability that a driver has a blood alcohol level higher than $0.05\%$? 

- P(A) = 0.38 (from part b)
- $P(A \cap B)$ = 0.08
$$P(B|A) = {P(A \cap B) \over P(A)}$$ 
$= \frac{0.08}{0.38}$ =  0.2105263 = 21.05%
```{r}
probablity_positivetest_higher <- 0.08/0.38
probablity_positivetest_higher
```

	
d) Suppose $A$ is the event that a driver has a positive test result and $B$ is the event that the driver has a blood alcohol level above $0.05\%$. Are $A$ and $B$ independent? Justify your answer. 

Formulas for independent events for A and B: 
$P(A \cap B) = P(A) \times  P(B)$

Also two events $A$ and $B$ are independent if $P(B|A) = P(B)$

We already know:
- P(A) = 0.38
- P($A \cap B$) = 0.08

So: P(B) = sum of alcohol level (both negative and positive) above $0.05\%$ = 0.08 + 0.02 = 0.1

```{r}
p_of_b <- 0.08 + 0.02
p_of_b

```

Verification using formula for independent: P($A \cap B$)  = P(A) $\times$ P(B) = 0.38 \times 0.1 = 0.038

```{r}
verify <- 0.38*0.1
verify
```

```{r}
P_A_intersect_B <- 0.08
P_A <- 0.38
P_B <- 0.1

if (P_A_intersect_B == P_A * P_B) {
  print("It is independent")
} else {
  print("It is NOT independent")
}
```

This means that P(A \cap B) is not same as P(B) since 0.038 is not same as 0.1 as $P(A \cap B) ≠ P(A) \times P(B)$ ***Therefore, not independent***
