---
title: "Homework Assignment 2"
author: ""
format:
  html:
    toc: true
    toc-depth: 3
    number-sections: true
    code-fold: false
execute:
  echo: true
  warning: false
  message: false
---

## March work day (10 pts)

R functions for common set operations (when you have *actual sets/vectors* of outcomes):

```{r, echo = TRUE, eval = FALSE}
x <- c(1, 2, 3)
y <- c(2, 3, 4, 5)
union(x, y)      # combined elements
intersect(x, y)  # shared elements
setdiff(x, y)    # elements in x but not y (order matters)
setequal(x, y)   # are the sets equal?
```

You have a vector of March 2026 dates (`march`), and vectors `weekends` and `springbreak`. Use R’s set operations to find the dates of **non‑vacation weekdays**.

```{r}
march <- 1:31
weekends <- c(1, 7, 8, 14, 15, 21, 22, 28, 29)
springbreak <- 16:20
```

## Two yellow balloons (20 pts)

**Problem :**  
There are 16 balloons: 5 yellow, 8 red, 3 green. Two darts pop two balloons (without replacement).

Let:

- $P(1_{yellow})$ = probability first balloon is yellow
- $P(2_{yellow})$ = probability second balloon is yellow

1. Given first is yellow, find $P(2_{yellow}\mid 1_{yellow})$.



2. Find $P(1_{yellow}\cap 2_{yellow})$ (both balloons yellow).



## Sequential system with A, B, C (15 pts)

A sequential system fails if **any** component fails. Components fail independently with:

- $p_A=0.02$
- $p_B=0.01$
- $p_C=0.03$

Find $P(\text{system fails})$.


## Problem parallel system (15 pts) 

A parallel system functions if **at least one** component functions. Components function **independently** with:

- $p_D=0.78$
- $p_E=0.75$
- $p_F=0.82$

Find $P(\text{parallel system functions})$.

## Spam detection using keywords problem (Bayes' Theorem) (20 pts)


An email service provider is designing a spam filter. From historical data, the company knows:

* **20%** of all incoming emails are spam.
* A specific keyword (for example, “FREE!!!”) appears in:

  * **60%** of spam emails
  * **5%** of legitimate (non-spam) emails

Suppose a new email arrives and it **contains this keyword**. What is the probability that the email is spam? 

## Two way contingency table (20 pts)

Suppose you are helping the police develop a new test for blood alcohol levels. The potential advantage of the test is that it can be used during a routine traffic stop, but neither you nor the police are sure how accurate the test is. By design, the test should be positive when blood alcohol level is above $0.05\%.$

The test has been used on a large number of subjects where a more expensive, less convenient test is known to give perfectly accurate results. The table below shows the proportions for the outcome of the new test and true blood alcohol status:

| Test result | Alcohol level > 0.05%  | Alcohol level <= 0.05% |
|-------------|------------------------|------------------------|
| Positive    | 0.08                   | 0.3                    |
| Negative    | 0.02                   | 0.6                    |
	
a) What is the probability that a randomly selected driver has both a positive test and a blood alcohol level higher than $0.05\%$? 
	
b) For a randomly selected driver, what is the probability that the test will be positive? 
	
c) Among drivers who have a positive test, what is the probability that a driver has a blood alcohol level higher than $0.05\%$? 
	
d) Suppose $A$ is the event that a driver has a positive test result and $B$ is the event that the driver has a blood alcohol level above $0.05\%$. Are $A$ and $B$ independent? Justify your answer. 
