---
title: "Homework Assignment 4"
author: "Anna Sehgal"
#output: pdf_document
format:
  html:
    toc: true
    toc-depth: 3
    number-sections: true
    code-fold: false
execute:
  echo: true
  warning: false
  message: false
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
set.seed(123)
```

## 1. The Poisson Distribution (40 pts Total)

**Concept:** Models the number of events occurring within a fixed interval of time or space.

### Scenario: Website Traffic

A website receives an average of **5 hits per minute**. Assume that the number of hits the website receives
follows Poisson distribution

1. Find the probability of receiving exactly 3 hits in a minute.
The probability of receiving exactly 3 hits in a minute

Given $\lambda = 5$

From notes from week 5 we know that, Probability Mass Function (P.M.F) is
$$ f(x;\lambda) = \frac{\lambda^{x} \cdot e^{-\lambda}}{x!}$$

Here x would be equal to th enumber of hits in a minute. Therfore $x = 3$

Substituting values: 
$$ f(x=3;\lambda=5) = \frac{5^{3} \cdot e^{-5}}{3!}$$
$$=> \frac{125 \cdot e^{-5}}{3 \cdot 2 \cdot 1}$$
$$=> \frac{125 \cdot e^{-5}}{6}$$

Value of $e^{-5}$ is:
```{r}
exp(-5)
```

Therefore $e^{-5} = 0.006737947$

$$=> \frac{125 \cdot 0.006737947}{6}$$
```{r}
threehitsprobability <- (125*0.006737947)/6
threehitsprobability
```

$$ f(x=3;\lambda=5) = 0.1403739$$

Therefore, the probability of receiving exactly 3 hits in a minute  is 0.1403739. That is 14.03739%.

Visual:
```{r}
lambda = 5
max_x = 30

#for plotting diff range
x <- 0:max_x
pmfx <- dpois(x, lambda)
poisx <- data.frame(x =x , pmfx =pmfx)
prob_3 <- dpois(3, lambda)
library(dplyr)
library(ggplot2)
poisx %>% ggplot(aes(x = x, y = pmfx, fill = x == 3)) + 
  geom_col(alpha = 0.7, width = 0.5) +
  scale_fill_manual(values = c("FALSE" = "blue", "TRUE" = "pink"), guide = "none") + labs(title = "Poisson Distribution (λ = 5)", subtitle = paste("P(X = 3) =", round(prob_3, 4)), x = "Number of Hits", y = "Probability") + theme_minimal()
```


2. Find the probability of receiving less than 10 hits in a minute.

Given xvals < 10 hits, $\lambda$ = 5

$$ f(x;\lambda) = \sum_{x=0}^{xvals} \frac{\lambda^{x} \cdot e^{-\lambda}}{xvals!}$$
P(xvals < 10) = P(x ≤ 9)

$$ f(x;\lambda) = \sum_{x=0}^{9} \frac{5^{xvals} \cdot e^{-5}}{xvals!}$$

```{r}
# it would be computed using ppois, not 1-ppois since it is not X ≥ value but instead X ≤ 9 (x> 10)
lamdba <- 5
tenhitprobablity <- ppois(9, lamdba) 
tenhitprobablity
```

Therefore, the probability of receiving less than 10 hits in a minute (x > 10) is 0.9681719, which is 96.81719%.


3. Simulation & Visualization

a.Generate a vector called hits with length 1000 that simulates the number of hits using rpois().

```{r}
lamdba <- 5
hits <- rpois(1000, lamdba)
hits
```

b.Create a histogram of this data with hist().

```{r}
library(ggplot2)
lambda = 5
hits <- rpois(1000, lambda)
hist(hits)
```

c.Add a vertical red line at the mean using abline().
```{r}
library(ggplot2)
lambda = 5
hits <- rpois(1000, lambda)
hist(hits)
#https://stackoverflow.com/questions/56875191/how-to-plot-simple-vertical-line-with-abline-in-r
meanhits <- mean(hits)
abline(v= meanhits, col="red", lwd=2)
```


## 2. The Normal Distribution (30 pts, 15 pts each)


### Scenario: A factory produces precision ball bearings. The target diameter is 25.00 mm. Due to machine variance, the diameters follow a Normal distribution with a mean ($\mu$) of 25.00 mm and a standard deviation ($\sigma$) of 0.02 mm.

a: Defect RatesBearings are considered defective if they are smaller than 24.95 mm or larger than 25.05 mm. Use pnorm() to calculate the total proportion of defective bearings.

Given:

$\mu$ = 25.00mm 

$\sigma$ = 0.02 mm

RatesBearings are defective if they x < 24.95mm or if x > 25.05mm.

Formula for Z from week notes for standard normal distribution :
$$Z := {X - \mu \over \sigma}$$
for the following: 
$$P(a < X < b) = P\left( {a - \mu \over \sigma} < Z < {b - \mu \over \sigma} \right) = \Phi\left( {b - \mu \over \sigma} \right) - \Phi\left( {a - \mu \over \sigma} \right).$$
Therefore: 
For X = 24.95 mm
Formula for Z from week notes for standard normal distribution :
$$Z = {24.95 mm - 25.00mm  \over 0.02 mm} = -2.5$$
```{r}
lower = (24.95-25.00)/0.02
lower
```

So, $P(X < -2.5)= P(Z < -2.5)$ for X = 24.95 mm


And for X = 25.05 mm
Formula for Z from week notes for standard normal distribution :
$$Z = {25.05 mm - 25.00mm  \over 0.02 mm} = 2.5$$
```{r}
upper = (25.05-25.00)/0.02
upper
```
So, $P(X > 2.5) = P(Z > 2.5)$ for X = 25.05 mm

Total proportion would be added in this case: So P(X <24.95 mm) + P(X>25.05 mm).
```{r}
mu <- 25 #mean
sigma <- 0.02 #sd
lowerp <- pnorm(q = 24.95, mean = mu, sd = sigma)
lowerp
```

```{r}
mu <- 25 #mean
sigma <- 0.02 #sd
upperp <- 1 - pnorm(q = 25.05 , mean = mu, sd = sigma) # due to upper
upperp
```

```{r}
lowerp <-  0.006209665
upperp <- 0.006209665
totaldefectp <- lowerp + upperp
totaldefectp
```

Therefore, the total proportion of defective bearings P(totaldefective) = 0.01241933 which is 1.241933%

b: Identifying Thresholds
The quality control team wants to flag the smallest 1% of bearings for manual inspection. Use qnorm() to find the diameter threshold for this bottom 1%.

R command:
```{r}
mu <- 25 #mean
sigma <- 0.02 #sd
thresholds <- qnorm(p = 0.01, mean = mu, sd = sigma)
thresholds
```
Based on the function, threshold for bottom 1% is 24.95347 mm.

Verification: Figure out the smallest 1% = 0.01 for P(X<x) =0.01

Formula for Z from week notes for standard normal distribution :
$$Z_{0.01} = {X - \mu \over \sigma}$$

Here X = thresholds

So,
$$Z_{0.01} = {thresholds - \mu \over \sigma}$$
plugging values:
$$Z_{0.01} = {24.95347 - 25 \over 0.02}$$
```{r}
zvalueonepercent <- (24.95347 - 25)/0.02
zvalueonepercent
```

Verification:
$$Z \cdot \sigma= X- \mu$$ 
$$Z \cdot \sigma + \mu= X$$

$$=> X = \mu + Z \sigma$$ 
$$=> X = 25 + (-2.33)*0.02$$ 
```{r}
Xvalue <- 25 + (-2.33)*0.02
Xvalue
```
Therefore, verified that 24.9534 mm is the diameter threshold for this bottom 1%.

---

## 3. The Exponential Distribution (10 pts each Total: 30 pts + Bonus 10 pts)

**Concept:** Models the time between events in a Poisson process.

### Scenario: Call Center Waiting Time

The time between calls at a help desk follows an exponential distribution with a rate  of **2 calls per hour**.
For the help desk, assume arrivals are a Poisson process.

a. Let $T$ be the waiting time (in hours) until the next ticket. What distribution does $T$ follow and what is its rate parameter?
Given that:

rate = $\lambda = 2$ calls per hour

X = T

T~Exp($\lambda$)

from Module 5 notes, exponential distribution:
$$f(x|\lambda) = \begin{cases} \lambda \exp{(-\lambda x)} & x \geq 0 \\ 0 & x < 0 \end{cases}$$
T~Exp(2)

Since x≥0 so $\lambda \exp(-\lambda x)$

Therefore: 
$$f(T|\lambda) = 2 \exp(-2T)$$


b. Compute the probability you wait **more than 45 minutes** for the next ticket.

P(45 minutes)= P(0.75 hours)

Let $T \sim \text{Exponential}(\lambda = 2)$. The cumulative distribution function (CDF) is derived from the probability density function (PDF):

PDF:  
$$f(T) = \lambda e^{-\lambda t}, \quad T \ge 0$$

CDF:  
$$F(T) = P(T \le t) = \int_0^T \lambda e^{-\lambda x} \, dx$$

$$
=>\int_0^T \lambda e^{-\lambda x} \, dx = \Big[- e^{-\lambda x} \Big]_0^T = -(e^{-\lambda T} - 1) = 1 - e^{-\lambda T}
$$

So the CDF is:  
$$F(T) = 1 - e^{-\lambda T}$$

Then the probability of waiting more than t is:
$$P(T > T) = 1 - F(T) = e^{-\lambda T}$$
$$P(T > 0.75) = 1 - F(T) = e^{-2 \cdot 0.75}$$
```{r}
lambda <- 2
t <- 0.75 
probt <-exp(-lambda * t)
probt
```

Therefore it is 0.2231302 for probability of waiting more than 45 minutes which is 0.75 hours.


c. Compute the median waiting time.
Median is the middle = 50% = 0.50
P(T≤wait) = 0.5
$$F(T) = 1 - e^{-\lambda T}$$
$$0.5 = 1 - e^{-\lambda T}$$
$$=> 1 - 0.5 = e^{-\lambda T}$$
$$=> 0.5 = e^{-\lambda T}$$
$\lambda$ = 2

So $-2 T = log(0.5)$

Then, $T = -log(0.5)/2$

```{r}
T <- -log(0.5)/2
T
```

Therefore, P(T>45 minutes) median is 0.3465736 hrs, which is 20.794416 minutes.


(Challenge Question) d. Memoryless property check: verify numerically that
   $P(T > s+t \mid T > s) = P(T > t)$ using s = 0.5, t = 0.3. 
   
   
   $$P(T > 0.5+0.3 \mid T > 0.5) = P(T > 0.3)$$
   
We know : Then the probability of waiting more than t is:
$$P(T > t) = 1 - F(t) = e^{-\lambda t}$$

So,
$$P(T>0.3) = e^{-2 \cdot 0.3}$$
```{r}
-2*0.3
```
$$P(T>0.3) = e^{-0.6}$$

and for 
   $$P(T > 0.5+0.3 \mid T > 0.5)$$
    $$=>P(T > 0.8 \mid T > 0.5)$$
$$=>P(T > 0.8)/P(T > 0.5)$$

from: 
$$P(T > t) = 1 - F(t) = e^{-\lambda t}$$
P(T > 0.8)= $e^{-2 \cdot 0.8}$ = 

```{r}
-2*0.8
```
P(T > 0.8)= $e^{-1.6}$ 

P(T > 0.5) = $e^{-2 \cdot 0.5}$ =
```{r}
-2*0.5
```
P(T > 0.5)= $e^{-1}$ 

$$=>e^{-1.6}/e^{-1}$$ 
$$=>e^{-0.6}$$ 

Therefore both sides are $e^{-0.6} = e^{-0.6}$. So:
```{r}
exp(-0.6)
```

Meaning that both sides are 0.5488116 = 0.5488116.
Therefore, verified.
